---
title: "Styling 'parallelPlot'"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Styling 'parallelPlot'}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7
)
```

```{r setup}
library(parallelPlot)
```

*This text was originally published on the [IFPEN forum](https://ifpen-discourse.appcollaboratif.fr/t/styling-parallelplot-or-scatterplotmatrix/163).*

Someone asked me for help styling a `parallelPlot` using the `cssRules` argument. In his case, he would like to hide the slider at the top of the chart.

```{r}
parallelPlot(iris)
```

Let's see how to find the solution following three steps.

## Step 1 - Find which `CSS` rule to use

When we open a `parallelPlot` from a R console, it's in fact a `HTML` page which is created.

> `HTML` provides the fundamental building blocks for structuring Web documents and apps.
> 
> `CSS` is used to describe the appearance of Web documents and apps.

[https://developer.mozilla.org/en-US/docs/Web](https://developer.mozilla.org/en-US/docs/Web)

To find which `CSS` rule to use, we can rely on a search engine. The main keyword to use is `css`. Searching "css hide element", the first result points to:

[https://developer.mozilla.org/en-US/docs/Web/CSS/visibility](https://developer.mozilla.org/en-US/docs/Web/CSS/visibility)

Thanks to the "Try it" section, the interesting option seems to be `visibility: hidden`. 

## Step 2 - Make some tests to determine where to apply the `CSS` rule

- download the file [mtcars.html](https://gitlab.com/drti/parallelplot/-/blob/master/test/mtcars.html) and open it with a web browser (firefox will be used here);

- place the mouse pointer hover the part of the graph you want to hide (the top slider) and right click to select 'Inspect';
  (note: when the inspector is open, as you move the mouse around the elements listed in the `HTML` pane, the corresponding elements are highlighted in the page)

- in the `HTML` pane, if this is not already the case, click the element which seems the highest in the hierarchy and which seems to correspond to the top slider; it should be `<g class="slider" transform="translate(10,25)">`;

- right click and select 'Attributes/Add attributes' (or double click the `>` which is at the end of the line) and type `visibility: hidden` as seen in **Step 1**. The slider should disappear.

## Step 3 - Write a `CSS` code to use as `cssRules` argument

> `CSS` is a rule-based language — you define the rules by specifying groups of styles that should be applied to particular elements or groups of elements on your web page.

To learn more about `CSS`, see [https://developer.mozilla.org/en/docs/Learn/CSS/First_steps](https://developer.mozilla.org/en/docs/Learn/CSS/First_steps)

The `cssRules` argument expects a list of `CSS` rules. A `CSS` rule has two parts:

- a `CSS` rule opens with a selector. 

  > In CSS, selectors are patterns used to match, or select, the elements you want to style.

  In our example, we want to select the `HTML` element `<g class="slider" transform="translate(10,25)">`. `g` stands for "group", so we wants to select a group with a `class` attribute equals to `slider`.

- associated to this selector, there are one or more declarations, which take the form of property and value pairs. In our example, we need only one declaration, we want to use `visibility: hidden`.

So, in a R console, we can try:

```{r}
parallelPlot(iris, cssRules = list("g" = "visibility: hidden"))
```

But doing so, we hide to many elements, the displayed graph appears empty, all white. We need to use a more discriminating selector by specifying the class of the targeted group. To specify a class, the operator is the dot (to see the list of existing operators: [https://developer.mozilla.org/en-US/docs/Web/CSS/CSS_selectors](https://developer.mozilla.org/en-US/docs/Web/CSS/CSS_selectors)).

In a R console, we can try:

```{r}
parallelPlot(iris, cssRules = list("g.slider" = "visibility: hidden"))
```

**Bingo!**

This time, only the slider is hidden, while the graph itself is clearly visible.
